---
title: "COVID-19 Growth in Virginia"
author: "Robert Link"
date: "4/1/2020"
output: 
  pdf_document: default
  rmarkdown::html_vignette: default
vignette: >
  %\VignetteIndexEntry{COVID-19 Growth in Virginia}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup}
suppressPackageStartupMessages(library(dplyr))
library(tidyr)
library(ggplot2)
library(CovMitigation)
library(ggthemes)
library(latex2exp)
```

## Assumptions

We have to make a slight approximation here.  The growth rate we are interested
in is the growth in active cases, which is affected by both the infection rate and
the recovery rate.  The data from the department of health only has new cases (which
grow a little more slowly) and cumulative cases (which grow a little faster).  For
this analysis we went with new cases.

## Growth observations

```{r growthrate, fig.width=7, fig.height=4}
pltdata <- 
  select(uvads_covid19, date, vaNewCases, ftest, ntest) %>%
  filter(vaNewCases > 0, ftest > 0)
t <- seq_along(pltdata$date) - 1  # start at zero
pltdata$t <- t

## Compute the testing fraction correction.
ftest_relative <- pltdata$ftest / pltdata$ftest[1]
ntest_relative <- pltdata$ntest / pltdata$ntest[1]
#fcorrect <- min(pltdata$ftest) / pltdata$ftest
pltdata$correctedNewCases <- pltdata$vaNewCases / ftest_relative / pltdata$vaNewCases[1]
pltdata$correctedCumCases <- cumsum(pltdata$correctedNewCases)

## Label positioning
lastidx <- length(t)
lastdate <- pltdata$date[lastidx]

## Create the benchmark data
tds <- c(3,4,5,6,7)
tdlabs <- paste('td=',tds)
tdlabstex <- sprintf('$t_d = %d$', tds)
gr <- lapply(tds, 
             function(td) {
               r <- log(2)/td
               exp(r*t)
             })
benchmarks <- as.data.frame(gr)
names(benchmarks) <- tdlabs
benchmarks$date <- t + pltdata$date[1]
benchmarks <- pivot_longer(benchmarks, -date, names_to = 'doubling time')
benchplt <- 
  ggplot(data=pltdata, aes(x=date)) + 
  geom_line(data=pltdata, mapping=aes(y=correctedCumCases), size=1.2) +
  geom_line(data=benchmarks, mapping=aes(y=value, group=`doubling time`), color='lightgrey') +
  scale_color_solarized() +
  scale_y_log10() +
  theme_bw()
for(i in seq_along(gr)) {
  benchplt <- benchplt + geom_label(x=lastdate, y=log10(gr[[i]][lastidx]), label=TeX(tdlabstex[i]))
}
suppressWarnings(print(benchplt))
```
