---
title: "Differences between old and new observations"
author: "Robert Link"
date: "4/21/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(CovMitigation)
library(vacovdata)

strtdate <- as.Date('2020-03-18')
normdate <- as.Date('2020-03-18')
normftest <- 1.94e-5

addstats <- function(dat, version) 
{
  ## Compute the testing fraction correction.
  normidx <- which(normdate == dat$date)
  message(version, ': normidx: ', normidx, '    ftest[normidx]: ', dat$ftest[normidx])
  #dat$ftest_relative <- dat$ftest / dat$ftest[normidx]
  dat$ftest_relative <- dat$ftest / normftest
  
  dat$correctedNewCases <- dat$vaNewCases / dat$ftest_relative 
  dat$correctedCumCases <- cumsum(dat$correctedNewCases)
  dat$rawCumCases <- cumsum(dat$vaNewCases)
  
  #dat <- filter(dat, date >= strtdate)
  relidx <- which(strtdate==dat$date)
  dat$correctedNewCasesRel <- dat$correctedNewCases / dat$correctedNewCases[relidx]
  dat$correctedCumCasesRel <- dat$correctedCumCases / dat$correctedCumCases[relidx]
  dat$rawCumCasesRel <- dat$rawCumCases / dat$rawCumCases[relidx]
  
  dat$version <- version
  
  dat
}

ctpdata <- 
  select(vacovdata, date, positiveIncrease, ftest, totalTestResultsIncrease) %>%
  rename(vaNewCases=positiveIncrease, ntest=totalTestResultsIncrease) %>%
  filter(ntest > 0) %>% 
  arrange(date) %>%
  addstats('CTP')

olddata <-
  select(uvads_covid19, date, vaNewCases, ftest, ntest) %>%
  filter(ntest > 0) %>%
  arrange(date) %>%
  addstats('OLD')

bothdata <- bind_rows(ctpdata, olddata)
```

## Summary

**The data from the COVID Tracking Project (CTP) appears to produce a dramatically
different testing-corrected growth rate than our home-grown dataset**.  These 
differences appear to be due to errors in the home-grown
dataset in the early part of the data (prior to about 
15 March).  However, with the CTP data the growth rate, corrected for testing growth,
is absurdly low (see below), suggesting that something is wrong with the correction
methodology.  Until we get that sorted out, we should probably show raw growth
without attempting to correct for testing growth.  Below we show that the daily
testing rate seems to have stabilized, making correction for testing rate less
important anyhow.


## Raw new cases differ a bit early on, then converge
```{r rawnewcases}
ggplot(data=bothdata, aes(x=date, y=vaNewCases, color=version)) + geom_line(size=1.2) + 
  scale_y_log10() +
  theme_bw()
```

Note also that the CTP dataset has cases earlier than the old dataset did.  

## This creates a small difference in raw cumulative cases
```{r rawcumcases}
ggplot(data=bothdata, aes(x=date, y=rawCumCases, color=version)) + geom_line(size=1.2) + 
  scale_y_log10() +
  theme_bw()
```

The slope in this log-scale plot is the growth rate uncorrected for growth in testing.  Note
that the growth rates, as well as the actual values, converge as we go later in time.
Note also that neither dataset is perfectly exponential; the slopes are decreasing
over time.

## The two datasets also have a different number of tests early on
```{r ntest}
orig_strt <- as.Date('2020-03-18')
ggplot(data=bothdata, aes(x=date, y=ntest, color=version)) + geom_line(size=1.2) +
  geom_vline(xintercept=orig_strt) +
  geom_smooth(linetype=2, se=FALSE) +
  geom_label(x=orig_strt, y=3000, label='Original normalization date', color='black') +
  theme_bw()
```

The correction factor is computed by taking the number of tests relative to the
number of tests on the normalization date.

Note also that this is not a log-scale plot, so the growth over most of March
is linear, not exponential.  Note also that the smoothed trend (dashed line)
seems to have leveled off since the beginning of April, indicating that daily
testing seems to have stabilized at about 2200 tests per day.

## Thus, the relative ftest factor is also a bit different
```{r ftest}
ggplot(data=bothdata, aes(x=date, y=ftest_relative, color=version)) + geom_line(size=1.2) + 
  theme_bw()
```

Corrected new cases are calculated by dividing raw new cases by this factor.

## This amplifies the difference in the corrected new cases

```{r cornewcases}
ggplot(data=bothdata, aes(x=date, y=correctedNewCases, color=version)) + geom_line(size=1.2) + 
  scale_y_log10() +
  theme_bw()
```

## This discrepancy produces substantially different corrected cumulative cases
```{r correctedcumcases}
ggplot(data=bothdata, aes(x=date, y=correctedCumCases, color=version)) + geom_line(size=1.2) + 
  scale_y_log10() +
  theme_bw()
```

## When we plot growth rates starting on 18 March, they look very different

```{r correctedcumcasesrel}
ggplot(data=filter(bothdata, date >= strtdate),
       aes(x=date, y=correctedCumCasesRel, color=version)) + geom_line(size=1.2) + 
  scale_y_log10() +
  theme_bw()
```

The CTP dataset looks like it shoots up dramatically in the first week, then
moderates, while our old dataset has a less dramatic initial rise, but a higher
long-term growth rate.

```{r dbltimectp}
gt2 <- filter(ctpdata, correctedCumCasesRel >= 2.0)
dbldate <- min(gt2$date)
dbltime <- dbldate - strtdate
```

So far, so good, but that line for the CTP data crosses 2 on `r dbldate`, a doubling
time of `r as.numeric(dbltime)` days, which seems wholly unbelievable.  This makes me 
think there is some kind of flaw in the way I'm computing the correction factor,
but I can't put my finger on what it is.



