---
title: "R Notebook"
output: html_notebook
---

```{r setup}
library(doParallel)
library(dplyr)
library(here)
library(metrosamp)
library(ggplot2)
library(tidyr)
library(CovMitigation)

registerDoParallel(8)

mkdensplt <- function(samps) {
  varnames <- colnames(samps)
  pltdata <- pivot_longer(as.data.frame(samps), everything(), names_to = 'parameter')
  pltdata$parameter <- factor(pltdata$parameter, levels=varnames, ordered=TRUE)
  ggplot(data=pltdata, aes(x=value)) + geom_density(fill='LightGrey', alpha=0.5) + facet_wrap(~parameter, scales='free') +
    theme_bw()
}

set.seed(867-5309)
```

```{r loaddata}
datadir <- here('analysis', 'mcmc-rslts', 'mcmc-2020-04-21')
warmup_files <- list.files(path=datadir, pattern='mcmc-.*\\.rds', full.names=TRUE)

mcmcruns <- lapply(warmup_files, readRDS)
allsamps <- getsamples(mcmcruns, 10000)
cruns <- metrosamp2coda(mcmcruns, 10000)
```

## Expectation value and credible intervals
```{r basicanalysis}
cat('Expectation values:\n')
print(signif(EV(mcmcruns), 2))
cat('\nCredible intervals:\n')
print(signif(CI(mcmcruns), 2))
cat('\nEffective samples:\n')
print(neff(cruns))
cat('\nRhat statistic:\n')
coda::gelman.diag(cruns)
```

## Effective Reproduction Number

```{r reff}
samps <- rsample(mcmcruns, 1000)
reffvals <- apply(samps, 1, calcreff)
cat('EV(R_e):\n', signif(mean(reffvals), 2), '\n')
cat('\nCI(R_e):\n')
signif(quantile(reffvals, c(0.025, 0.975)), 2)
```

## Density plots

```{r densplots}
# pltdata <- pivot_longer(as.data.frame(allsamps), everything(), names_to = 'parameter')
# ggplot(data=pltdata, aes(x=value)) + geom_density(fill='LightGrey', alpha=0.5) + facet_wrap(~parameter, scales='free') +
#   theme_bw()
mkdensplt(allsamps)
```

```{r sampleruns}
Nsamp <- 100
parmsamps <- rsample(mcmcruns, Nsamp)
runparms <- function(parms, scenname) {
  modparms <- parms[c('A0', 'D0', 'T0')]
  raw <- run_scenario(270, as.list(modparms), scenname)
  totalUVA <- 
    group_by(raw, time, scenario) %>% 
    summarise(newSymptoCases = sum(newSympto * marketFraction)) %>%
    ungroup()
  dt <- totalUVA$time + parms['day_zero']
  ## align the results to integer days
  dtout <- floor(dt)
  interpcases <- approx(dt, totalUVA$newSymptoCases, dtout, rule=2)
  dateout <- dtout + as.Date('2020-01-01')
  tibble(scenario=totalUVA$scenario, date=dateout, newSymptoCases=interpcases$y)
}

sample_runs <- 
  lapply(seq(1,Nsamp), 
         function(i) {
           parms <- parmsamps[i,]
           runparms(parms, as.character(i))
         }) %>%
  bind_rows()

ev_run <- runparms(EV(mcmcruns), 'EV')
```

```{r ribbonplot}
strtdate <- as.Date('2020-04-01')
enddate <- as.Date('2020-09-01')
ci_by_date <- 
  filter(sample_runs, date >= strtdate, date < enddate) %>%
  #mutate(time = as.numeric(date-strtdate)) %>%
  group_by(date) %>%
  summarise(nslo = quantile(newSymptoCases, 0.1), nshi = quantile(newSymptoCases, 0.9)) %>%
  ungroup()

pltev <- filter(ev_run, date >= strtdate, date < enddate)

rplt <- 
  ggplot(data=pltev, aes(x=date)) + 
  geom_line(aes(y=newSymptoCases), size=1.2) +
  geom_ribbon(data=ci_by_date, aes(ymin=nslo, ymax=nshi), alpha=0.5) +
  theme_bw()
print(rplt)
```

## Investigating the bimodality

```{r bimodality, fig.width=10}
sampslp <- getsamples(mcmcruns, 10000, TRUE)
sampslp <- cbind(sampslp$samples, sampslp$lp)
colnames(sampslp)[ncol(sampslp)] <- 'Lpost'
bvals <- sampslp[,'b']
losamps <- sampslp[bvals < 65,]
hisamps <- sampslp[bvals > 71,]

mkdensplt(losamps)
mkdensplt(hisamps)

cat('losamps:\n')
quantile(losamps[,'Lpost'], c(0.025, 0.5, 0.975))
cat('hisamps:\n')
quantile(hisamps[,'Lpost'], c(0.025, 0.5, 0.975))
```

Astonishingly, these two subsamples seem to be about equally good.  What's going
on with the ones in the middle?

```{r midvals}
midsamps <- sampslp[bvals > 65 & bvals < 71,]
mkdensplt(midsamps)
cat('midsamps:\n')
quantile(midsamps[,'Lpost'], c(0.025, 0.5, 0.975))
```

So, the best samples from that mid-range are about as good as the median sample
from either the low samples or the high samples.  Why?  What's different about 
these samples?

```{r cicmp}
cimat <- function(mat, p=c(0.025, 0.975)) {
  apply(mat, 2, quantile, probs=p)
}

ml <- t(cimat(losamps))
colnames(ml) <- paste(colnames(ml), '(lo)')
mh <- t(cimat(hisamps))
colnames(mh) <- paste(colnames(mh), '(hi)')
mm <- t(cimat(midsamps))
colnames(mm) <- paste(colnames(mm), '(mid)')
signif(cbind(ml,mh,mm), 4)
```

