---
title: "Model Results Summary"
author: "Robert Link"
date: "4/17/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(CovMitigation)
library(dplyr)
library(ggplot2)
library(doParallel)

registerDoParallel(8)

mparms <-
  c(T0_hi = 6.23326057949715, T0 = 9.54243712017808, D0 = 6.73682352612848, 
    A0 = 3.77160066817391, I0 = 11.3521753597075, Ts = 3.96825, day_zero = 43.0690562665808, 
    b = 91.1258634441669)

pltcounties <-
  c("FairfaxCounty", "ArlingtonCounty", "PrinceWilliamCounty", 
    "HenricoCounty", "Charlottesvillecity", "AlbemarleCounty")

daterng <- as.Date(c('2020-04-01', '2020-09-01'))

disp <- function(plt) {
  suppressWarnings(print(plt))
}
```

## Introduction

These are the outputs of the results of the UVA infection model for the MAP 
(maximum _a posteriori_) model parameters using calibration data through 2020-04-15.
These results represent our best guess at the modal trajectory of the infection 
in the Commonwealth of Virginia through the end of the summer; however, as yet
there are a wide range of parameters that fit the data adequately, and so there
is a great deal of uncertainty surrounding these projections.

This analysis was generated with version `r packageVersion('CovMitigation')` of
the model.

## Comparison of model to data.

This figure shows a comparison between the model projections and observed data
for several representative localities.  Fairfax, Arlington, Prince William, and
Henrico counties continue to show a higher growth rate than the rest of the 
Commonwealth.  The doubling time fit by the model for these counties was 
`r signif(mparms['T0_hi'], 2)` days, while the doubling time else where was fit
at `r signif(mparms['T0'], 2)` days.

```{r modobs, echo=FALSE, fig.width=8}
moplt <- plt_modobs(mparms, 'MAP fit', pltcounties)
disp(moplt + 
        scale_color_manual(values = 'MidnightBlue') + 
        facet_wrap(~locality, scales='free_y') +
        guides(color=FALSE))
```

## Model projections

These are the projections for daily new symptomatic cases over time, aggregated
across the state.

```{r prjstate, echo=FALSE, fig.width=8}
pltprj_state <- plt_projections(mparms, 'MAP Scenario')
disp(pltprj_state + 
        scale_color_manual(values='MidnightBlue') +
        guides(color=FALSE) +
        ylab('Daily new symptomatic cases (total)') +
        xlim(daterng))
```

Similar projections, adjusted for our historical market share by county

```{r prjuva, echo=FALSE, fig.width=8}
pltprj_state <- plt_projections(mparms, 'MAP Scenario', marketadjust = TRUE)
disp(pltprj_state + 
        scale_color_manual(values='MidnightBlue') +
        guides(color=FALSE) +
        ylab('Daily new symptomatic cases (market share adjusted)')+
        xlim(daterng))
```


Finally, this plot shows the cumulative fraction of the population infected, 
aggregated across the state, over time.  It represents the fraction over time of
people who have or have had the disease.
```{r prjcumul, echo=FALSE, fig.width=8}
pltprj_cumulfrac <- plt_projections(mparms, 'MAP Scenario', what='PopCumulFrac')
disp(pltprj_cumulfrac +
        scale_color_manual(values='MidnightBlue') +
        guides(color=FALSE) + 
        xlim(daterng) +
        ylab('Cumulative infection fraction'))
```