---
title: "Reopening Scenarios"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
library(dplyr)
library(nleqslv)
library(ggplot2)
library(doParallel)
library(CovMitigation)
registerDoParallel(8)

agg_scen <- function(scendata) {
  agg <- 
    group_by(scendata, time, scenario) %>%
    summarise(I=sum(I), Is=sum(Is), R=sum(R),
              newCases=sum(newCases),
              newSympto=sum(newSympto)) %>%
    ungroup()
  agg$totalpop <- 8517685
  mutate(agg, date=as.Date('2020-01-01')+time, Itot=I+Is, 
         cumInfect=Itot+R, cumIfrac=cumInfect/totalpop) %>%
    filter(time-floor(time) < 1e-8)
}
```

## Introduction

We wish to explore the consequences of "reopening" the Commonwealth of Virginia.
The governor has suggested that the COVID-19 mitigation policies might be lifted
after two consecutive weeks of declines in new cases [1].  What are the likely
consequences of such a policy?  Does it risk creating a resurgence in the infection?

## Analysis
Begin with the MAP parameter set.  Run it and aggregate to the state level.

```{r base_scen}
pmap <-
  c(T0_hi = 5.69195605740336, T0 = 7.81804244536788, D0 = 6.73879591630125, 
    A0 = 3.52224288149668, I0 = 2.56384762262623, Ts = 3.98902851913746, 
    day_zero = 29.0878000968141, b = 59.7317732176905)
map_scen <- run_parms(pmap, scenario_name = 'Base scenario')
map_scen_agg <- agg_scen(map_scen)
max_sympto_i <- which.max(map_scen_agg$newSympto)
max_sympto_date <- map_scen_agg$date[max_sympto_i] 
```

This scenario reaches its maximum daily new symptomatic cases on
`r max_sympto_date`.  We will take this as 
indicative of the date at which the observed cases also peak.  One scenario for
reopening the economy that has been discussed is to wait until two weeks of consistent
decline in new case observations.  Because of the random factors that affect the number
of cases actually measured, that could be quite a bit more than two weeks after the
peak, but let's assume they do some kind of smoothing, and the reopening happens
two weeks after the peak.

To analyze this scenario, we first have to decide what reopening is likely to do
to our model parameters.  We will assume that it returns the infection transmissibility
to a value characteristic of a disease with $R_0= 2.5$ and solve for the 
change in $\beta$ that makes that happen.  Since $R = \beta/\gamma$, and we aren't
changing $\gamma$, it follows that $\frac{\beta_r}{\beta} = \frac{R_r}{R}$.  

The implied $R$ for our MAP (maximum _a posteriori_) parameters is about 
`r signif(calcr0(pmap), 2)`.  Since
that isn't a huge reduction from our presumptive value of 2.5 for the unmitigated
disease, we'll also run a $\beta$ for an $R_0$ of 5.  We don't really
expect it to be that high, but that should allow us to bracket the possibilities.

With that in mind, the two schedules for the model $\beta$ parmeter are:

```{r reopenbeta}
targr0 <- 2.5
peaklag <- 14
r0 <- calcr0(pmap)
beta_fac <- targr0/r0
beta_time <- map_scen_agg$time[max_sympto_i] + peaklag
beta_sched <- data.frame(time=c(0,beta_time), value=c(1,beta_fac))
beta_sched_hi <- data.frame(time=c(0, beta_time), value=c(1,2*beta_fac))
beta_sched
beta_sched_hi
```
Each of these tables shows the values of $\beta$ relative to its value at the 
beginning of the simulation, as a function of time.  The value of $\beta$ is 
constant in between entries in the table, so in these cases $\beta$ remains at
its original value until $t = `r beta_time`$, when it gets multiplied by the
factor shown in the second row of the table.  

Now we can run this same scenario, but with the new values of $\beta$ starting on
`r max_sympto_date + peaklag`.  Note that as currently implemented the model will
increase the statewide transmissibility and the high-growth-rate transmissibility
by the same factor.  We probably want to set them both to the same assumed reference
level, but that involves adding an option to include a second beta factor specific
to the high growth rate, which we don't have yet.  We're going to assume that by
the time the state as a whole is two weeks past its peak, the high-growth regions
are far enough past their peak that the inaccuracy is insignificant.

```{r reopenscen}
reopenparms <- c(as.list(pmap[1:6]), list(beta_schedule=beta_sched))
tstrt <- pmap['day_zero']
tnext <- ceiling(tstrt)
tvals <- c(tstrt, seq(tnext, 273))
reopen_scen <- run_scenario(tvals, reopenparms, 'Reopening')
reopen_scen_agg <- agg_scen(reopen_scen)

hireopenparms <- c(as.list(pmap[1:6]), list(beta_schedule=beta_sched_hi))
hireopen_scen_agg <- agg_scen(run_scenario(tvals, hireopenparms, 'Reopening high-spread'))
```


```{r cmpplt}
pltdata <- bind_rows(map_scen_agg, reopen_scen_agg, hireopen_scen_agg)
strtdate <- max_sympto_date
enddate <- max(pltdata$date)
plt <- ggplot(pltdata, aes(x=date, y=newSympto, color=scenario)) + geom_line(size=1.2) +
  scale_color_brewer(type='qual') +
  xlim(c(strtdate,enddate)) +
  theme_bw() 
print(plt)
```

## Conclusion

Even if we assume that the transmissibility of the disease increases dramatically
when the mitigation policies are lifted, the rate of new infections continues to
drop.  The reason for this, in this particular model, is that even with mitigation
policies in place, the rate of new infections is large enough at the peak that even
once new infections begin to decline, the overshoot is large enough the population
reaches the herd immunity threshold even for pessimistic assumptions about the 
unmitigated disease.

It's worth keeping in mind some of the assumptions underlying these results.  The
biggest is that people who recover from the disease acquire immunity to it, which 
is currently a subject of vigorous investigation.  More generally, the model used
here has a variety of assumptions including constancy of coefficients, persistence of
the difference in growth patterns between the high and low growth regions, and 
independence of growth in different counties.  This last assumption, in particular,
might be expected to hold while mitigation policies are in effect and travel is
limited, but it might not be valid for a scenario with more relaxed policies.

Also assumed in these results is that authorities can reliably determine the peak
of the infection from observed confirmed cases.  Observed data is noisy, and it 
is entirely possible to be fooled by a false peak, though a two-week confirmation
period makes this less likely.  More subtly, changes in testing procedures could 
produce the appearance of a peak whilst the infection is still spreading.  This could
happen, for example, if testing becomes less targeted, which could offset the 
increase in infection prevalance and cause the fraction
of observed positive tests to decline.  If no correction were made for the change
in procedure, the decline in positive results could be misinterpreted as a peak
in the infection.


## References

[1] https://www.richmond.com/news/virginia/not-there-yet-northam-outlines-future-plans-for-reopening-the-state/article_e76696ef-e361-5629-95d1-b6c35a91922c.html
